% !TEX root = ../thesis-example.tex
%
\pdfbookmark[0]{Acknowledgement}{Acknowledgement}
\chapter*{Acknowledgement}
\label{sec:acknowledgement}
\vspace*{-10mm}
First and foremost, I would like to express my sincere gratitude to my promotor, Prof. Dr. Archisman Ghosh, for his guidance, support, and insightful feedback throughout this project. I am especially thankful for his flexibility and willingness to accommodate my schedule, which enabled me to work effectively under my own constraints.

I am particularly grateful to my supervisor, Cezary Turski, whose mentorship, patience, and consistent encouragement were instrumental throughout every phase of this thesis. His feedback, technical guidance, and responsiveness were key in helping me navigate both conceptual challenges and practical implementation. I deeply appreciate the time and effort he invested in my progress, and his understanding in adjusting to my working rhythm.

I am also grateful to the Ghent Gravity Group for fostering a welcoming, inclusive, and intellectually stimulating environment, one that made even a master's student feel part of the broader research effort.

A special thanks goes to Prof. Marcelle Soares-Santos (University of Zurich) for her early advice and for introducing me to the \texttt{BUZZARD} mock catalog, which shaped an important part of this thesis.

I gratefully acknowledge the use of computational resources provided by the LIGO Laboratory computing clusters, which enabled the analyses performed in this work. Special appreciation is due to the developers and maintainers of the \texttt{gwcosmo} pipeline, the \texttt{GWSim} package, and the galaxy catalogs used in this study, including \texttt{GLADE+} and \texttt{BUZZARD}, whose open resources made this research possible.

On a personal note, I am deeply grateful to my family for their unwavering support throughout this journey. I owe a special debt of gratitude to my parents, whose love, encouragement, and sacrifices have made everything possible. I would also like to thank my friends for their timely distractions and unwavering companionship, which served as crucial reminders that there is, in fact, life beyond academic jargon and late-night analyses.

Finally, I would like to thank everyone who contributed, directly or indirectly, to the successful completion of this thesis. Your support has meant more than words can convey.


\vspace{2em}
\noindent Muhammad Khuzaifa Naveed
