\chapter{Conclusion}
\label{chap:conclusion}

The emergence of \ac{GW} astronomy has opened a new frontier in observational cosmology. This thesis has explored the potential of using dark sirens, gravitational-wave events without electromagnetic counterparts, for the inference of the Hubble constant, $H_0$. Specifically, we examined whether refining the galaxy catalog used in the redshift prior, by restricting it to its brightest subsets, to trace the large-scale structure, can improve the cosmological constraints from these events.

\section{Summary of Findings}

We implemented a hierarchical Bayesian inference framework using the \texttt{gwcosmo} pipeline, which combines \ac{GW} luminosity distance posteriors with redshift priors constructed from galaxy catalogs. The galaxy redshift prior was derived using the \texttt{GLADE+} catalog and modified by applying brightness cuts to prioritize the most luminous galaxies in the $K$-band. These subsets, denoted \texttt{GLADEPXX}, were hypothesized to trace large-scale structure more efficiently due to their association with massive halos.

The analysis demonstrated that:
\vspace{-1em}
\begin{itemize}
    \item Brightness-ranked catalogs improve the redshift prior, leading to tighter posteriors on $H_0$ by reducing low-likelihood host candidates in the \ac{GW} localization volume.
    \vspace{-1em}
    \item Moderate pruning yields improvement in $H_0$ precision as compared to using the full catalog, without introducing measurable bias. This improvement is attributed to the enhanced clustering signal from retaining mostly central, high-luminosity galaxies.
    \vspace{-1em}
    \item Aggressive pruning (e.g., \texttt{GLADEP10}) results in degraded constraints and larger uncertainties, as the catalog no longer adequately traces the large-scale structure.
    \vspace{-1em}
    \item A lower bound of approximately 30\% of the brightest galaxies emerges as a practical pruning limit, in the best-case scenario where the brightest galaxies are strictly the central-most galaxies. Below this threshold, significant information about the underlying matter distribution is lost, particularly from central galaxies, increasing the risk of cosmological bias.
    \vspace{-1em}
    \item Mock data challenges using the \texttt{BUZZARD} catalog support the robustness of the brightness-based pruning strategy but also emphasize the importance of catalog completeness, redshift depth, and optimal percentile thresholds. The 30\% structural limit appears consistent across lower magnitude cuts, suggesting that future work should avoid overly aggressive pruning or instead develop hybrid approaches combining full and pruned catalogs, depending on the redshift regime.
\end{itemize}

\section{Limitations}

This work, while comprehensive, has several limitations:
\vspace{-1em}
\begin{itemize}
    \item The number of \ac{GW} events with high \ac{SNR} and good localization remains small, limiting the statistical power of our results.
    \vspace{-1em}
    \item Catalog incompleteness, particularly at high redshifts, introduces uncertainties that are only partially mitigated by brightness cuts.
    \vspace{-1em}
    \item Due to time constraints and several technical challenges encountered with the \texttt{BUZZARD} mock catalog, we were unable to perform a full end-to-end bias quantification from pruning dim galaxies. Nevertheless, partial analyses and theoretical considerations support the proposed 30\% threshold as a conservative lower bound.
    \vspace{-1em}
    \item While a complete end-to-end \acf{MDC} framework has been developed, a few components still require minor refinements and consistency checks before full deployment in future analyses.
    %\vspace{-1em}
    %\item The analysis assumes a fixed $\Lambda$CDM cosmology throughout, exploring extended models may reveal different degeneracies or sensitivities in dark siren inference.
\end{itemize}

\section{Outlook and Future Work}

The outlook for dark siren cosmology is highly promising. With the advent of next-generation \ac{GW} detectors such as Cosmic Explorer~\citep{Evans:CE}, Einstein Telescope~\citep{Abac:ET} and LISA~\citep{LISA:2024hlh}, and deeper galaxy surveys like LSST~\citep{ivezic2019lsst} and Euclid~\citep{mellier2024euclid}, both the number of detected events and the completeness of host catalogs are expected to improve significantly.

Future extensions of this work could include:
\vspace{-1em}
\begin{itemize}
    \item Getting a quantitative estimate of the bias introduced by the brightness cuts, with the devised end-to-end \ac{MDC} framework.
    \vspace{-1em}
    \item Testing for potential biases introduced by brightness cuts using larger and more realistic mock datasets.
    \vspace{-1em}
    \item Developing adaptive redshift priors that vary with localization volume depth, combining full catalogs at low redshift with bright subsets at high redshift.
    \vspace{-1em}
    \item Integrating clustering information or cross-correlations with large-scale structure to improve redshift inference~\citep{afroz2024prospect}.
\end{itemize}

Ultimately, the resolution of the Hubble tension will require a convergence of multiple independent probes. As shown in this thesis, dark sirens, when carefully analyzed, offer a robust and independent route to measuring $H_0$. Furthermore, the use of bright galaxies as tracers of large-scale structure provides a promising avenue for improving the precision of cosmological constraints from \ac{GW} observations. This methodology will play a growing role in the era of multi-messenger cosmology.
